---
title: "DataScience_1Assignment"
author: "Gary Leander"
date: "17/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```


# Data Science with R: Assignment Task

## 1. BASICS

### 1.1. Variables in  R

What is the value of  " B " after running these command?																													    [2]


```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t

```

The value of B is : `r round(B,2)`

What type of vector do you get in each of the following?																														   [4]
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```

```{r}
vectype1<-typeof(vec_1)
vectype2<-typeof(vec_2)
vectype3<-typeof(vec_3)
vectype4<-typeof(vec_4)
```

vec_1 is a `r vectype1` type of vector.

vec_2 is a `r vectype2` type of vector.

vec_3 is a `r vectype3` type of vector.

vec_4 is a `r vectype4` type of vector.


### 1.2. Conditional subsetting

Consider the following vector:

```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

vec_1<-vec
vec_1<-c(na.omit(vec))
```
`r vec_1`

vec_1 is formed by checking where in the vector "vec" there are missing values in the form "NA" and all those values are omitted through the "na.omit()" function.

#### 1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code   												     [3]

```{r}
x<-c(which(vec %in% NA))

```
`r x`

The code checks at which position NA appears in the original vector "vec" through the which() function and uses the positions as values for the new vector "x"


#### 1.2.3. Write a code to determine how many missing values are in vec.																													      [2]
```{r}
AmtMissing<-sum(is.na(vec))
```

There are `r AmtMissing` missing values in "vec"


#### 1.2.4. Write a R code to create a data frame which contains details of 3 students as follows:	   															  		[4]

- Name
- Surname
- Age
- Gender

```{r}
df<- data.frame("Name"=c("Name1","Name2","Name3"),"Surname"=c("Surname1","Surname2","Surname3"),"Age"=c(1,2,3),"Gender"=c("Male","Female","Female"))
```



## 2. Functions in R

### 2.1. Create  a function which returns a mean of an input vector and explain your code

```{r}
#x<-c()
meanVal <-function(x){
  a<-sum(x)
  b<-length(x)
  return(a/b)
  
}
x<-c(1,2,3,4)
meanVal(x)
```
the function takes a vector as its input, sums up the vector it was given, determines the length of the vector and returns the calculated mean (sum of vector divided by length of vector)

### 2.2. Create a function which take  no inputs , it reads  a Titanic .csv file and returns its content and explain your code. 								  [3]

```{r}
readxprint<- function(){
  df <- read.csv('train.csv', sep = ",")
  df
}

readxprint()

```

the readxprint function reads the train csv file separated by a "," and assigns it to df it then returns the contents of the train csv file.


## 3. Conditionals and Loops

#### 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse.														     																																						  [8]

```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE","MALE","MALE","FEMALE","MALE","MALE","FEMALE","MALE","MALE","MALE")
```

```{r}
chngMxF<-function(g){
  newg<-c(ifelse(g=="MALE",1,ifelse(g=="FEMALE",2,10)))
  newg
}
chngMxF(g)
```
`r chngMxF(g)`


3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
```


3.2.1. Use the "cbind " function in R to create one dataset and assign it to df. 																						  [2]

```{r}
df<-cbind(x,y)
df
```

3.2.2. Convert df from 3.2.1 into a data.frame   																																		  [2]

```{r}
df_new<-data.frame(df)
df_new

```

3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F")  in 3.2.1														[3]

```{r}
males<-df_new %>% filter(y=='M') %>% count()
females<-df_new %>% filter(y=='F') %>% count()

```

Number of males are: `r males`
Number of females are: `r females`


3.2.4. Create a bar plot using ggplot to visualize your results in 3.2.3																										[5]
```{r}
df_new %>% group_by(y) %>% count() %>%
   mutate(y=as.factor(y)) %>%
  ggplot(aes(x=y , y=n))+
  geom_col()+labs(x="Gender",y="Number of Males & Females", fill= "Gender",title = "Males vs Females")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

```

3.2.5. Use mutate function to add sum of x per "M" and "F" and plot the results using sum of x per "M" and "F".								     [5]
```{r}
xS<-sum(x)

```





